### 2021-04-27 papers/thesis/full/ch1-introduction/introduction.tex
Addressed more of SO comments

In this thesis we expand upon this knowledge by providing frameworks for similar sparse spectral methods for solving PDEs on other multidimensional domains (notably including the disk-slice and trapezium in 2D, and the spherical cap surface in 3D) that also yield matrices that are what is defined as \enquote{banded-block-banded}. We take inspiration from the work established for the triangle \cite{olver2019triangle} and the unit disk \cite{vasil2016tensor}, which can  be seen as special cases of the disk-slice and trapezium in the framework we present here. The spherical cap work serves to lay a foundation for using spherical caps and spherical bands as elements in a spectral element method for solving PDEs on the whole sphere, as an alternative to the Spherical Harmonic transform approach that the European Centre for Medium-range Weather Forecasts (ECMWF) use in their weather and climate model \cite{cheong2006dynamical}.

Spherical Harmonics are of course a long-established and famous group of functions defined on the surface of a sphere with certain useful properties (for example, they are orthogonal to each other on the unit sphere and form a basis for expanding functions defined on the sphere) and as a result are widely used for many applications involving PDEs \cite{}\bstodo{cite applications of SHs}. Notably, they are used as basis functions for the spectral transform method that makes up part of the model in the Integrated Forecasting System (IFS), which is used by ECMWF for their forecasts \cite{wedi2013fast}. While the whole sphere spectral method using the Spherical Harmonics has been successful for numerous years \cite{williamson2007evolution}, there is a drawback in the parallel scalability bottleneck that arises from the global spectral transform, which is expected to inhibit future performance of the IFS \cite{ecmwf2020scalability, wedi2013fast, diamantakisecmwf}.

>>>>

In this thesis we expand upon this knowledge by providing frameworks for similar sparse spectral methods for solving PDEs on other multidimensional domains (notably including the disk-slice and trapezium in 2D, and the spherical cap surface in 3D) that also yield matrices that are what is defined as \enquote{banded-block-banded}. We take inspiration from the work established for the triangle \cite{olver2019triangle} and the unit disk \cite{vasil2016tensor}, which can  be seen as special cases of the disk-slice and trapezium in the framework we present here. The spherical cap work serves to lay a foundation for using spherical caps and spherical bands as elements in a spectral element method for solving PDEs on the whole sphere, as an alternative to the Spherical Harmonic transform approach that the European Centre for Medium-range Weather Forecasts (ECMWF) use in their weather and climate model \cite{cheong2006dynamical}. There are also potential applications in physics, where

Spherical Harmonics are of course a long-established and famous group of functions defined on the surface of a sphere with certain useful properties (for example, they are orthogonal to each other on the unit sphere and form a basis for expanding functions defined on the sphere) and as a result are widely used for many applications including in computer graphics (e.g. \cite{moon2008efficient, sloan2013efficient}), physics (e.g. \cite{vasil2019tensor}), quantum theory \cite{varshalovich1988quantum} and geosciences \bstodo{cite applications of SHs}. Notably, they are also used as basis functions for the spectral transform method that makes up part of the model in the Integrated Forecasting System (IFS), which is used by ECMWF for their forecasts \cite{wedi2013fast}. While the whole sphere spectral method using the Spherical Harmonics has been successful for numerous years \cite{williamson2007evolution}, there is a drawback in the parallel scalability bottleneck that arises from the global spectral transform, which is expected to inhibit future performance of the IFS \cite{ecmwf2020scalability, wedi2013fast, diamantakisecmwf}.
